Some Graphical Realizations of Two-Row Specht Modules of Iwahori--Hecke Algebras of the Symmetric Group

### 2019-07-31 tex/Drafts/Draft-3.tex
General restructuring, conjecture section

  We will prove the following theorem on irreducibility of $M_{2n + r}^r$.
  \begingroup
  \hypersetup{linkcolor=\thmcolor} % no orange theorem label
  \begin{customthm}{\ref{Irreducibility Theorem}}
    Suppose $e > n$ and $S^{(n+r,n)}$ is irreducible.
    Then $M_{2n + r}^r$ is irreducible.
  \end{customthm}
  \endgroup
  Note that the representations $M_{0 + r}^r$ and $S^{(r)}$ are both isomorphic to the sign representation.
  This and Theorem \ref{Irreducibility Theorem} are suggestive;
  in fact, we will prove the following.
  \begingroup
  \hypersetup{linkcolor=\thmcolor}
  \begin{customthm}{\ref{Correspondence Theorem}}
    Suppose $e > n + r + 1$.
    Then, $M_{2n + r}^r \simeq S^{(n+r,n)}$.
  \end{customthm}
  \endgroup
  Lastly, we will conjecture that $M_{2n + r}^r \simeq S^{(n+r,n)}$ for all $e$, and support this with heuristic data.

@@

  \vspace{5pt}
  In Section \ref{Crossingless Matchings Section}, we begin by proving Theorem \ref{Irreducibility Theorem}.
  Following this, we prove the existence of a particular filtration with factors given by other crossingless matchings representations;
  using irreducibility, this becomes a composition series.
  This combined with an inductive argument and the branching of Section \ref{Specht Modules Section} allow us to prove Theorem \ref{Correspondence Theorem}.
  Last, we finish the section by proving more general statements concerning sign subrepresentations of $M$, which contribute to conjectures later in the paper.{\color{orange} Fill in with specific information later.}

@@

\section{Sign subrepresentations}
{\color{magenta} Smooth out the beginning of this section to work with the end of last section.}
In the previous section, we proved $M_{2n+r}^r\simeq S^{(n+r,n)}$ under certain conditions of $e$. In this proof, we used the fact that, given our conditions on $e$, we only projected down to cases with no sign subrepresentation.

@@

\subsection{Conjectures on sign subrepresentations}\label{Conjecture Section}
Recall that $K := K_{2n + r}^r$ is the direct sum of all copies of the sign representation in $M$.
We will begin with a promising characterization of the socle of some $M_{2n + r}^r$.
\begin{proposition}\label{Preimage of K}
  Fix some $M := M_{2n + r}^r$.
  Suppose there is some natural number $n'$ with $n' < n$ such that $K_{2n' + r}^r \neq 0$.
  Let $\pi:M_{2n + r}^r \twoheadrightarrow M_{2n' + r}^r$ be the linear map
  \[
    \pi := (1 + T_1)(1 + T_3)\cdots(1 + T_{n-n'}).
  \]
  Then, $\pi^{-1}(K)$ contains any proper subrepresentations of $M$.
\end{proposition}
\begin{proof}
  It suffices to show that any vector $w \in M - \pi^{-1}(K)$ is cyclic;
  then, any subrepresentation containing $w$ also contains all of $M$, implying the proposition.

  In fact, since $\pi(w) \notin K$, there is some $(1 + T_i) \in \SH$ such that $(1 + T_i)\pi(w) \neq 0$.
  Since $\ima (1 + T_i)$ is irreducible, we are then able to generate the image of a basis element, and hence generate a basis element in $M_{2n + r}^r$ by analogous argument to \ref{Irreducibility Theorem}. {\color{orange} justify soon tomorrow.}
  {\color{orange} Sketch: subrepresentation $W \supsetneq \pi^{-1}(K)$ maps down to proper subrepresentation containing $K$, a contradiction.}
\end{proof}

If $\pi^{-1}(K)$ is itself a subrepresentation of $M$, then we will have identified the radical of $M_{2n + r}^r$.
Observing decomposition matrices for $S_{(n+r,n)}$ for small $n$, we note that $S^{(n+r,n)}$ appears to have composition series of length at-most two by \cite{Mathas-book};
the same appears to be true of $M_{2n + r}^r$ by Figures \ref{Composition series e=3} through \ref{Composition series e=6}.
This motivates our first conjecture:
\begin{conjecture}\label{Preimage subrepresentation}
  In Proposition \ref{Preimage of K}, $\pi^{-1}(K)$ is an irreducible subrepresentation of $M$.
\end{conjecture}

We can posit more sturcutre than this;
by observing Figure \ref{Empirical Kernel}, we note that the kernel is nontrivial in more cases than $e = n + r + 1$, and appears to follow a pattern: 
\begin{conjecture}\label{Sharp kernel}
  Suppose $p = \infty$.
  Then, $K \subset M_{2n + r}^r$ is nontrivial if and only if $e \mid n + r + 1$, in which case it is one-dimensional.
\end{conjecture} 
Then, assuming that Conjectures \ref{Preimage subrepresentation} and \ref{Sharp kernel} are true, we reach the following corollary.
\begin{corollary}
  $M := M_{2n + r}^r$ has a composition series given by \begin{equation}\label{Conjecture filtration}
    0 \subset \pi^{-1}(K) \subset M.\end{equation}
\end{corollary}
\begin{proof}
  By \ref{Sharp kernel}, we have the filtration \eqref{Conjecture filtration}.
  By \ref{Preimage subrepresentation}, the first factor is irreducible, and by \ref{Preimage of K}, the second factor is irreducible.
\end{proof}

This characterizes much of the structure of $M_{2n + r}^r$.
Due to this structure, the heuristics listed above, their isomorphism in the generic case, and the equality of their dimension, we pose the following conjecture.
\begin{conjecture}
  For any $e,n,r$, we have $M_{2n + r}^r \simeq S^{(n+r,n)}$.
\end{conjecture}

@@

\iffalse
\begin{proposition}
  Denote the composition factor $M_{2n + r}^r / h^{-1}(K_{2n + r}^r)$ by $U_{2n + r}^r$.
  Then, there exist some naturals $m,s$ satisfying $2m + s = 2n + r$ and $m + s > n + r$ such that the following is an isomorphism of $\SH$-modules \[h^{-1}\prn{K_{2n + r}^r} \simeq U_{2m + s}^s\].
  \qed
\end{proposition}
\begin{proposition}
  For the same $m,s$ as above, we have the following composition series of Specht modules:
  \[
    0 \longrightarrow D^{(m+s,m)} \longrightarrow S^{(n+r,n)} \longrightarrow D^{(n+r,n)} \longrightarrow 0.
  \]
\end{proposition}
\begin{proposition}
  $M_{2n + r}^r \simeq S^{(n+r,n)}$ and $U_{2m +s}^s \simeq D^{(m+s,m)}$.
\end{proposition}
\fi

>>>>

  We will prove the following theorem on irreducibility of $M_{2n + r}^r$.
  \begingroup
  \hypersetup{linkcolor=\thmcolor} % no orange theorem label
  \begin{customthm}{\ref{Irreducibility Theorem}}
    Suppose $e > n$ and $S^{(n+r,n)}$ is irreducible.
    Then $M_{2n + r}^r$ is irreducible.
  \end{customthm}
  \endgroup
  Note that the representations $M_{0 + r}^r$ and $S^{(r)}$ are both isomorphic to the sign representation.
  This and Theorem \ref{Irreducibility Theorem} are suggestive;
  in fact, we will prove the following.
  \begingroup
  \hypersetup{linkcolor=\thmcolor}
  \begin{customthm}{\ref{Correspondence Theorem}}
    Suppose $e > n + r + 1$.
    Then, $M_{2n + r}^r \simeq S^{(n+r,n)}$.
  \end{customthm}
  \endgroup
  Each of these are powerful characterizations of the module $M_{2n + r}^r$ in the case that it and $S^{(n+r,n)}$ are irrducible.
  We will prove the following theorem which characterizes $M_{2n + r}^r$ in the reducible case:
  \begingroup
  \hypersetup{linkcolor=\thmcolor}
  \begin{customthm}{\ref{kernel existence}}
    Suppose $e = n + r + 1$.
    Then, $M_{2n + r}^r$ contains a subrepresentation isomorphic to the sign representation. 
  \end{customthm}
  \endgroup
  We will go on to prove Corrollary \ref{n+r+1 composition series}, which specifies a composition series for $M_{2n + r}^r$ for the case $e = n + r + 1$.

@@

  \vspace{5pt}
  In Section \ref{Crossingless Matchings Section}, we begin by proving Theorem \ref{Irreducibility Theorem}.
  Following this, we prove the existence of a particular filtration with factors given by other crossingless matchings representations;
  using irreducibility, this becomes a composition series.
  This combined with an inductive argument and the branching of Section \ref{Specht Modules Section} allow us to prove Theorem \ref{Correspondence Theorem}.
  Last, we finish the section by proving more general statements concerning sign subrepresentations of $M$, which contribute to conjectures later in the paper.

  \vspace{5pt}

@@

  \section{Sign subrepresentations}\label{Sign section}
{\color{magenta} Smooth out the beginning of this section to work with the end of last section.}
In the previous section, we proved $M_{2n+r}^r\simeq S^{(n+r,n)}$ under certain conditions of $e$. In this proof, we used the fact that, given our conditions on $e$, we only projected down to cases with no sign subrepresentation.

@@

Due to the argument present in Theorem \ref{Irreducibility Theorem}, sign subrepresentations provide a surprisingly strong characterization of some $M_{2n + r}^r$ as follows.
\begin{proposition}\label{Preimage of K}
  Fix some $M := M_{2n + r}^r$.
  Suppose there is some natural number $n'$ with $n' \leq n$ such that $K_{2n' + r}^r \neq 0$.
  Let $\pi:M_{2n + r}^r \twoheadrightarrow M_{2n' + r}^r$ be the linear map
  \[
    \pi := \begin{cases}
      (1 + T_1)(1 + T_3)\cdots(1 + T_{n-n'}) & \text{if } n < n,'\\
      \id{M} & \text{if } n = n',
    \end{cases}
  \]
  where $\id M$ is the identity map on $M$.
  Then, $\pi^{-1}(K)$ contains any proper subrepresentations of $M$.
\end{proposition}
\begin{proof}
  It suffices to show that any vector $w \in M - \pi^{-1}(K)$ is cyclic;
  then, any subrepresentation containing $w$ also contains all of $M$, implying the proposition.

  In fact, since $\pi(w) \notin K$, there is some $(1 + T_i) \in \SH$ such that $(1 + T_i)\pi(w) \neq 0$.
  Note that $e > (n'-1) + r + 1$, so $M_{2(n-1) + r}^r$ is irreducible;
  then, we may use diagram \eqref{Irreducibility commutative diagram} and an analogous argument to Theorem \ref{Irreducibility Theorem} to argue that $(1 + T_i)\pi(w)$ is cyclic, and hence $w$ is cyclic.
\end{proof}
\begin{corollary}\label{n+r+1 composition series}
  Suppose $e = n + r + 1$.
  Then $M_{2n + r}^r$ has a composition series given by
  \[
    0 \subset K \subset M.
  \]
\end{corollary}

This is suggestive;
the preimage $\pi^{-1}(K)$ may be integral in characterizing the representation $M_{2n + r}^r$ in other reducible cases.
The following section poses several conjectures involving this structure.

\subsection{Conjectures on sign subrepresentations}\label{Conjecture Section}
If $\pi^{-1}(K)$ is itself a subrepresentation of $M$, then we will have identified the radical of $M_{2n + r}^r$ via Propostion \ref{Preimage of K}.
Observing decomposition matrices for $S_{(n+r,n)}$ for small $n$, we note that $S^{(n+r,n)}$ appears to have composition series of length at-most two by \cite{Mathas-book};
the same appears to be true of $M_{2n + r}^r$ by Figures \ref{Composition series e=3} through \ref{Composition series e=6}.
This motivates our first conjecture:
\begin{conjecture}\label{Preimage subrepresentation}
  In Proposition \ref{Preimage of K}, $\pi^{-1}(K)$ is an irreducible subrepresentation of $M$.
\end{conjecture}

We can posit more sturcutre than this;
by observing Figure \ref{Empirical Kernel}, we note that the kernel is nontrivial in more cases than $e = n + r + 1$, and appears to follow a pattern: 
\begin{conjecture}\label{Sharp kernel}
  Suppose $p = \infty$.
  Then, $K \subset M_{2n + r}^r$ is nontrivial if and only if $e \mid n + r + 1$ and $e < n$. in which case it is one-dimensional.
\end{conjecture} 
Then, assuming that Conjectures \ref{Preimage subrepresentation} and \ref{Sharp kernel} are true, we reach the following corollary.
\begin{corollary}
  Suppose $e,n,r$ are such that $M_{2n + r}^r$ is irreducible and $e \mid n' + r + 1$ for some $e < n' < n$.
  Then, $M := M_{2n + r}^r$ has a composition series given by \begin{equation}\label{Conjecture filtration}
    0 \subset \pi^{-1}(K) \subset M.\end{equation}
\end{corollary}
\begin{proof}
  By \ref{Sharp kernel}, we have the filtration \eqref{Conjecture filtration}.
  By \ref{Preimage subrepresentation}, the first factor is irreducible, and by \ref{Preimage of K}, the second factor is irreducible.
\end{proof}

These conjectures hope to characterize much of the structure of $M_{2n + r}^r$.
Due to this structure, the heuristics listed above, their isomorphism in the $e > n + r + 1$ case, and the equality of their dimension, we pose the following conjecture.
\begin{conjecture}
  For any $e,n,r$, we have $M_{2n + r}^r \simeq S^{(n+r,n)}$.
\end{conjecture}

@@
